Network Configuration Synthesis with Abstract Topologies \vspace{1cm}

### 2017-03-13 paper/abstraction/paper.tex
more edits, shrinking

\newcommand{\Topo}[1]{\ensuremath{{\mathrm{topo}}(#1)}}
\newcommand{\Pref}{\ensuremath{\mathrm{pref}}}

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=2.5in]{figures/example}
  \caption{An example data center network.}
  \label{fig:example}
  \vspace{-1em}
\end{figure}

\begin{figure}[t!]
  \centering
  \includegraphics[width=2.5in]{figures/example2}
  \caption{A modified version of the example in Figure~\ref{fig:example}.}
  \label{fig:example2}
  \vspace{-1em}
\end{figure}

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/templates}
  \caption{Idealized small configuration component for the data center spine based on templates from a cloud provider.}
  \label{fig:templates}
  \vspace{-1em}
\end{figure}

@@

To simplify network configuration, operators are adopting a template-based approach~\cite{hatch,thwack}. Instead of authoring a configuration per router, operators author a template per role.
%
A {\em role} is a specific function that is served by one or more routers.
%\footnote{If a role is played by devices from multiple vendors, there is one template per role-vendor combination because templates use vendor configuration languages.}
%
For example, the network in Figure~\ref{fig:example} might have five roles: spine, global aggregator, global ToR, local aggregator, and local ToR.
%
Figure~\ref{fig:templates} shows an example of what a small component of a template for the spine role in the two data centers might look like. The template has parameters for various aspects of the configuration (\EG, neighbor list, local prefixes) and are compiled to low-level device configurations by instantiating the parameters using the network topology and a database of network information.

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/example3}
  \vspace{-2em}
  \caption{An abstraction for the network in Figure~\ref{fig:example}.}
  \label{fig:example3}
  \vspace{-1em}
\end{figure}

@@

Our topology abstractions can also capture concrete topologies; there will be a one-to-one correspondence between abstract and concrete nodes (and edges). This ability enables operators to define complex networks in which they expect some (e.g., legacy) parts of the network to not evolve and others to evolve.
%It also means that our compiler subsumes those that operate only over concrete topologies.

@@

The second line introduces a prefix \emph{template} variable $\CD{\$GP}$. It says that traffic for each global prefix associated with the variable should follow a path that ends at its corresponding destination router in the $\CD{TG}$ role. As we will see in \S\ref{sec:language}, constraints like $\CD{\End(TG)}$ are just syntactic sugar for regular expressions describing network paths. The second line has a similar policy for local prefixes.
%
The final rule matches all other IP prefix destinations and allows traffic to follow a path that leaves the data center, ending at some outside location, through neighbors $\CD{N1}$ or $\CD{N2}$ with a preference for leaving through $\CD{N1}$. The $\CD{\Prefer}$~symbol indicates that traffic should satisfy the constraint on the left and resort to the backup (right) only when that is not possible due to network failures.%

Next we can capture constraint (4) that traffic for local prefixes must stay within in the data center:
%
\begin{code}
\Define Local =
    \$LP \Path \Always(\In)
\end{code}
\noindent%
%
The Local policy adds the $\CD{\Always(\In)}$ constraint for each local prefix described by the template variable $\CD{\$LP}$. This constraint ensures that traffic follows a path that matches an internal location (\IE, in the data center) at each hop of the path.
%
The constraint to prevent ``valleys" (5) is written as:
%
\begin{code}
\Define NoValley =
    \True \Path \Novalley(\{TG,TL\},\{AG,AL\},\{S\})
\end{code}
\noindent%
%
This policy applies to all traffic and prevents valley paths by adding the $\Novalley$ constraint with arguments corresponding to each level in the data center.
%
Constraint (7) to prevent transit traffic between neighbors is expressed as:
%
\begin{code}
\Define Peer = \{N1,N2\}
\Define NoTransit =
    \True \Path !(\Enter(Peer) & \Exit(Peer))
\end{code}
\noindent%
%
We define a $\CD{Peer}$ as $\CD{N1}$ or $\CD{N2}$ and disallow paths where traffic both enters and exits the data center through a peer.

@@

\begin{code}
Routing & Local & NoTransit &
NoValley & \Agg(PG, \In -> \Out)
\end{code}
\noindent
This instructs \sysname to satisfy the conjunction of all the constraints. It also declares that we want to perform route aggregation, as covering prefix PG, for global prefixes at the border of the data center (\IE, along any edge that connects the inside of the data center to the outside).

@@

\begin{figure}[t]\small
  \begin{minipage}[t]{\linewidth}
  \vspace*{-1\baselineskip}
  %
  \[ \begin{array}{rclr}
     d       &\in& \textit{Integers} \\
     x       &\in& \textit{Variables} \\
     pol     &::=& p_1, \dots, p_n & \text{Policy} \\
     p       &::=& t \hspace{.3em} \Path \hspace{.3em} r_1 \Prefer \dots \Prefer r_m \BNFALT a & \text{Path Preferences}  \\
     t       &::=& \$x \BNFALT d.d.d.d/[d..d] \hspace{1.5em} & \text{Predicate} \\
     a      &::=& \mathit{agg}(t, r_1 \rightarrow r_2) & \text{Aggregation} \\
     r       &::=& l \BNFALT \emptyset \BNFALT \CD{in} \BNFALT \CD{out} \BNFALT r_1 \cup r_2 \BNFALT & \text{Regular Path} \\
             &   & r_1 \cap r_2 \BNFALT r_1 \cdot r_2 \BNFALT \NOT r \BNFALT r^* \\
  \end{array} \]%

  \end{minipage}
  \vspace{-.3em}
  \caption{\sysname core syntax.}
  \label{fig:syntax}
  %\vspace{-0.8em}
\end{figure}%

@@

\begin{figure*}[t!]
  \begin{minipage}[t]{.4\linewidth}
  \hdr{Policy Automata}{}
  \vspace{1em}
    \begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=2cm,minimum size=.5cm]
      \state{0}{$0$}{              }{}
      \state{1}{$1$}{right of=0}{accepting}
      \transition{0}{0}{out}{loop above}
      \transition{0}{1}{N1}{}
      \transition{1}{1}{in}{loop above}
      \node at (1,-1){Rank 1 DFA};
    \end{tikzpicture}
    ~~~~~
    \begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=2cm,minimum size=.5cm]
      \state{0}{$0$}{              }{}
      \state{1}{$1$}{right of=0}{accepting}
      \transition{0}{0}{out}{loop above}
      \transition{0}{1}{N2}{}
      \transition{1}{1}{in}{loop above}
      \node at (1,-1){Rank 2 DFA};
    \end{tikzpicture}%
  \end{minipage}
  %
  ~~~~~~
  %
  \begin{minipage}[t]{.3\linewidth}
    \hdr{Concrete Topology}{}
    \vspace{1em}
    \includegraphics[width=.7\columnwidth]{figures/topology-con}
  \end{minipage}%
  %
  \begin{minipage}[t]{.3\linewidth}
    \hdr{Abstract Topology}{}
    \vspace{1em}
    \includegraphics[width=.57\columnwidth]{figures/topology-abs}
  \end{minipage}%

  \vspace{1em}
  \begin{minipage}[t]{.6\linewidth}
    \hdr{Concrete Product Graph}{}
    \includegraphics[width=.9\columnwidth]{figures/pg-con}
  \end{minipage}%
  \begin{minipage}[t]{.4\linewidth}
  \hdr{Abstract Product Graph}{}
    \includegraphics[width=.88\columnwidth]{figures/pg-abs}

@@

  \hrulefill
  \vspace*{.4em}%

  \caption{Product Graph construction for policy $\True ~\Path~ \CD{\Exit(\text{N1}} ~\Prefer~ \CD{\text{N2})}$.}
  \label{fig:example-compilation}
  %\vspace{-1em}
\end{figure*}

@@

%After converting the routing policy to the regular expression based language,
We are now ready to generate the PG. Intuitively, the PG captures topology and routing constraints by ``intersecting" the finite automata associated with the policy regular expressions with the graph structure of the topology.
%The PG compactly encodes all paths that protocol messages should be passed along through the network in order to guarantee that the BGP-speaking devices find policy-compliant paths dynamically.
%
%Although \sysname policies describe the routes that traffic should take through the network, network protocols (including BGP) typically disseminate router information in the opposite direction. For example, a router Y will tell its neighbor X that it has a route to the destination meaning that X can send traffic through Y. For this reason, we start by compiling
%
Because there can be a separate routing policy for each predicate matching traffic, we construct one PG for each predicate.

For each regular path constraint $r_i$ from $r_1 ~\Prefer~ .. ~\Prefer~ r_k$, we construct a DFA for the reverse of $r_i$. A DFA for $r_i$ is defined as a tuple ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma$ is the set of topology locations, $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
The PG is a tuple ($G'$, $start$, \Pref) where $G' = (V',E')$ has
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon V' \times V'$,
a unique starting vertex $start$,
and a preference function $\Pref \colon V' \rightarrow 2^{\set{1, \dots, j}}$, mapping nodes in the PG to a set of path ranks.

The PG is constructed by adding an edge from state $m = (l_m, q_{m_1}, \dots, q_{m_k})$ to $n = (l_n, q_{n_1}, \dots, q_{n_k})$ whenever $\sigma_i(q_{m_i}, l_n) = q_{n_i}$ for each $i$ and $(l_m,l_n) \in E$.
%
We add edges from the $\mathit{start}$ node to any $m = (l, q_{m_1}, \dots, q_{m_k})$ when $\sigma_i(q_{0_i}, l) = q_{m_i}$ for each $i$.
%
The preference function $\Pref(m)$ denotes the rank of paths through the PG ending at node $m$ and is defined as $\Pref(m) = \set{i~\vert~q_{m_i} \in F_i}$.
%
Finally, we write $\Topo{m} = l$ to extract the topology location from a PG node, when $m = (l, q_{m_1}, \dots, q_{m_k}) \in V'$.

@@

  \hrule
  \vspace*{1em}
  \caption{Abstract k-disjoint path analysis inference rules.}
  \label{fig:inference-rules}
  \vspace{-1em}
\end{figure*}

@@

Figure~\ref{fig:inference-rules} displays the collection of rules used to infer facts about disjoint paths.
Each rule is read from bottom to top. The label on the bottom left is a known fact. We use $L$ to represent a rule that is parametric over the label ($S$ or $A$). The labels on other nodes correspond to facts learned after applying the inference rule. The adjacent box shows the conditions that must be valid, given the abstraction constraints, for the rule to apply. 
%Validity ensures that the condition holds for all concrete instantiations of the topology.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figures/analysis}
  \end{center}
  \vspace{-1.4em}
  \caption{Abstract disjoint path analysis for global prefixes. \label{fig:compilation-times}}
  \label{fig:example-inference}
  \vspace{0em}
\end{figure}

@@

The abstract disjoint path analysis starts from a fixed source location $src$ and repeatedly tries to apply every inference rule from Figure~\ref{fig:inference-rules} until it reaches a fixed point. The algorithm applies an inference rule when the rule's condition is valid given the abstract topology constraints. Because the inference rules may continue to yield larger and larger symbolic expressions, we make the following observation to ensure termination: for any invariant learned of the form $L(j,k)$, it is sound to instead infer $L(j',k')$ if $j' \leq j$ and $k' \leq k$. Therefore, for each inference $L(j,k)$ we minimize the symbolic expressions for $j$ and $k$ subject to the topology constraints using the optimizing SMT solver $\nu$Z~\cite{z3opt}. Facts learned with $j=0$ or $k=0$ are discarded.

@@

@@

At a high level, the translation from the PG representation to per-device templates that run the distributed BGP protocol involves the following steps: (1) Every BGP message is tagged to record the state of the product graph as messages are passed between routers. The tags allow BGP to only search for valid paths by dropping messages that do not correspond to any edge in the PG. (2) To ensure that BGP always converges to the \emph{most preferred} paths in the network, we must rank route advertisements for each router by giving them a preference such that, under those preferences, the network as a whole satisfies the \sysname policy's end-to-end network preferences.

We introduce a simple, vendor-independent BGP configuration language called \mbgp and describe a compilation function that uses the inferred preference ordering to generate concrete configurations from a concrete topology, or templates from an abstract topology. Templates can then be soundly instantiated for a concrete network by replacing instances of abstract neighbors with the union of all concrete neighbors under the inverse homomorphism. We prove that concretization and compilation commute. By modifying compilation slightly we can further ensure that every configuration depends only on its immediate neighbors. This guarantees that any change made to the concrete topology will result in a minimal number of changes to the configurations.

@@

The BGP routing protocol standard allows for the use of \emph{community tags}, 32-bit integer tags that can be arbitrarily attached to, or removed from routing advertisements. 
Community tags serve as a simple form of \emph{history}, and operators routinely use such tags to implement policy. For example, operators might record where an advertisment enters the network using community tags and then conditionally block the export of the advertisment later to other peers to prevent becoming a transit point between peers.

@@

For example, in the concrete PG in Figure~\ref{fig:example-compilation}, router S1 appears in two different PG nodes: $(S1,1,0)$ and $(S1,0,1)$. These nodes have two different peers that can send them messages in the PG: $(N1,0,0)$ and $(N2,0,0)$. Therefore, S1 will allow an advertisement from peer $N1$ and will also allow an advertisement from peer $N2$. If S1 uses the path advertised by neighbor N1, then it will add the tag $(1,0)$ to the advertisement before sending this to its neighbors A1 and A2. If S1 uses a path advertised from neighbor N2, then it will add the tag $(0,1)$ instead. Similarly, router A1 appears in two separate PG nodes. It will not block advertisements from S1 if they either have the $(1,0)$ tag or the $(0,1)$ tag attached. In either case, it will not modify the tag before exporting the route to its neighbors.

%If A1 chooses a path with tag $(1,0)$ attached, then it will export the path to its neighbors without modifying the tag. If A1 chooses a path with tag $(0,1)$ attached, it will also export the path to its neighbors without modifying the tag.

@@

  \vspace{1em}
  \hrulefill%
  %\vspace{1em}%

  \caption{mBGP syntax (left), and compilation from product graphs (right).}
  \label{fig:concretization}
  \vspace{-.8em}
\end{figure*}%

@@

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=\columnwidth]{figures/configs}
  \end{center}
  \vspace{-1em}
  \caption{Spine template and concrete configurations (left), and evolution-friendly templates (right).}
  \label{fig:bgp-configs}
\end{figure}

@@

The idea then is to search for such a device-local preference function for each router that totally orders advertisements from different neighbors (possibly with different tags). For example, to satisfy the policy that leaving through N1 is preferred to leaving through N2, S1 should prefer a message it hears from N1 over N2. Similarly, A1 should prefer routes tagged with $(1,0)$ over those with $(0,1)$. Intuitively, S1 should prefer a message from N1 because it results in an accepting state for automata 1, which indicates a more preferred path. Further, it can result in downstream routers such as A1 obtaining a route that is accepting for automata 1 as well, \IE $(A1,1,0)$ instead of $(A1,0,1)$.

@@

\begin{defn}
Let $m \geq_{rank} n$ be a relation over PG vertices that holds iff $\min (\Pref(m)) \geq \min (\Pref(n))$ or $\Pref(n) = \emptyset$. Intuitively $m \geq_{rank} n$ means that paths ending at node $m$ are no better than paths ending at $n$.
%
\end{defn}
\noindent
%

@@

Figure~\ref{fig:bgp-configs} (left) shows part of the generated \mbgp configuration for spine routers for both the concrete and abstract policies. For brevity, we using the symbol ($*$) to denote the set of all neighbors and omit tags when irrelevant.
%
For prefix $\True$ ($0.0.0.0/[0..32]$), the spine routers will match advertisements from peer N1 and N2. The match for N1 is preferred since it has a higher BGP local-preference attribute (110). If an advertisement from N1 is available the spine uses this and attaches the community tag $(1,0)$ before sending the route to all its peers ($*$). If an advertisement is only available from the backup N2, then it attaches the tag $(0,1)$ instead.
%
The template configuration matches any global prefix $\CD{\$GP}$ from any internal peer and re-advertises the route to all of its peers. For any local prefix it will allow an advertisement from any internal peer, and re-advertise the route to only other internal peers. The concrete configurations for S1 and S2 obtained from compilation for the concerete PG from Figure~\ref{fig:example-compilation} have a similar structure for each local and global prefix where local routes are reflected downward, while global routes are advertised to all peers.

@@

Suppose an operator wants to expand the concrete data center from Figure~\ref{fig:example} by adding an additional ToR router to the TG role. Per the network routing policy, the new ToR router will advertise any owned prefixes provided by looking up {\small \$GP} in $\Gamma$. Because the new topology matches the abstraction, the compiled templates will remain the same. However, in the spine router configurations, the match on the global prefix template variable {\small \$GP} must be expanded when concretizing the template to include the new prefixes added by the ToR. Unfortunately, this small change to the topology results in a change to every single spine router configuration.

@@

\sysname is implemented in approximately 9500 lines of F\# code. The compiler accepts both concrete and abstract topologies and will generate configurations for open-source Quagga~\cite{quagga} routers by default. The fault-tolerance analysis uses $\nu$Z~\cite{z3opt} to both test validity and minimize variables subject to the topology constraints. Since the analysis typically calls the SMT solver many times with relatively small optimization problems, we use a timeout of 200ms.

>>>>

\newcommand{\Topo}[1]{\ensuremath{{\mathrm{topo}}(#1)}}
\newcommand{\Pref}{\ensuremath{\mathrm{rank}}}

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=2.5in]{figures/example}
  \vspace{-.8em}
  \caption{An example data center network.}
  \label{fig:example}
  \vspace{-.8em}
\end{figure}

\begin{figure}[t!]
  \centering
  \includegraphics[width=2.5in]{figures/example2}
  \vspace{-.8em}
  \caption{A modified version of the example in Figure~\ref{fig:example}.}
  \label{fig:example2}
  \vspace{-.8em}
\end{figure}

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/templates}
  \caption{Idealized small configuration component for the data center spine based on templates from a cloud provider.}
  \label{fig:templates}
  \vspace{-.8em}
\end{figure}

@@

To simplify network configuration, operators are adopting a template-based approach~\cite{hatch,thwack}. Instead of authoring a configuration per router, operators author a template per role.
%
A {\em role} is a specific function that is served by one or more routers.
%\footnote{If a role is played by devices from multiple vendors, there is one template per role-vendor combination because templates use vendor configuration languages.}
%
For example, the network in Figure~\ref{fig:example} might have five roles: spine, global aggregator, global ToR, local aggregator, and local ToR.
%
Figure~\ref{fig:templates} shows an example of what a small component of a template for the spine role in the two data centers might look like. The template has parameters for various aspects of the configuration (\EG, neighbor list, local prefixes) and is compiled to low-level device configurations by instantiating the parameters using the network topology and a database of network information.

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/example3}
  \vspace{-2.2em}
  \caption{An abstraction for the network in Figure~\ref{fig:example}.}
  \label{fig:example3}
  \vspace{-.8em}
\end{figure}

@@

Our topology abstractions can also capture concrete topologies; there will be a one-to-one correspondence between abstract and concrete nodes/edges. This allows operators to define complex networks in which some (e.g., legacy) parts of the network can not evolve while others can.
%It also means that our compiler subsumes those that operate only over concrete topologies.

@@

The second line introduces a prefix \emph{template} variable $\CD{\$GP}$. Template variables represent multiple instances of a rule for different concrete prefixes that can be provided by an external source (\EG, a database). The line says that traffic for each global prefix associated with the variable should follow a path that ends at its corresponding destination router in the $\CD{TG}$ role. As we will see in \S\ref{sec:language}, constraints like $\CD{\End(TG)}$ are just syntactic sugar for regular expressions describing network paths. The second line has a similar policy for local prefixes.
%
The final rule matches all other IP prefix destinations and allows traffic to follow a path that leaves the data center, ending at some external role ($\CD{\Out}$), through neighbors $\CD{N1}$ or $\CD{N2}$ with a preference for leaving through $\CD{N1}$. The $\CD{\Prefer}$~symbol indicates that traffic should satisfy the constraint on the left and resort to the backup (right) only when that is not possible due to network failures.%

Next we can capture constraint (4) that traffic for local prefixes must stay within in the data center:
%
\begin{code}
\Define Local =
    \$LP \Path \Always(\In)
\end{code}
\noindent%
%
The Local policy adds the $\CD{\Always(\In)}$ constraint for each local prefix described by the template variable $\CD{\$LP}$. This constraint ensures that traffic follows a path that matches an internal role (\IE, in the data center) at each hop of the path.
%
The constraint to prevent ``valleys" (5) is written as:
%
\begin{code}
\Define NoValley =
    \True \Path \Novalley(\{TG,TL\},\{AG,AL\},\{S\})
\end{code}
\noindent%
%
This policy applies to all traffic and prevents valley paths by adding the $\Novalley$ constraint with arguments corresponding to each level in the data center.
%
Constraint (7) to prevent transit traffic between neighbors is expressed as:
%
\begin{code}
\Define Peer = \{N1,N2\}
\Define NoTransit =
    \True \Path !(\Enter(Peer) & \Exit(Peer))
\end{code}
\noindent%
%
We define a $\CD{Peer}$ as $\CD{N1}$ or $\CD{N2}$ and disallow paths where traffic both enters and exits the data center through a peer.

@@

\begin{code}
Routing & Local & NoTransit &
NoValley & \Agg(GP_AGG, \In -> \Out)
\end{code}
\noindent
This instructs \sysname to satisfy the conjunction of all the constraints. It also declares that we want to perform route aggregation, with covering prefix {\small GP\_AGG}, for global prefixes at the border of the data center (\IE, along any edge that connects the inside of the data center to the outside). In this case, {\small GP\_AGG} is declared as a concrete prefix rather than a template because a single aggregate prefix will be used to summarize all less-specific prefixes.

@@

\begin{figure}[t]\small
  \begin{minipage}[t]{\linewidth}
  \vspace*{-1\baselineskip}
  %
  \[ \begin{array}{rclr}
     d       &\in& \textit{Integers} \\
     x       &\in& \textit{Variables} \\
     l       &\in& \textit{Topology Locations} \\
     pol     &::=& p_1, \dots, p_n & \text{Policy} \\
     p       &::=& t \hspace{.3em} \Path \hspace{.3em} r_1 \Prefer \dots \Prefer r_m \BNFALT a & \text{Path Preferences}  \\
     t       &::=& \$x \BNFALT d.d.d.d/[d..d] \hspace{1.5em} & \text{Predicate} \\
     a      &::=& \mathit{agg}(t, r_1 \rightarrow r_2) & \text{Aggregation} \\
     r       &::=& l \BNFALT \emptyset \BNFALT \CD{in} \BNFALT \CD{out} \BNFALT r_1 \cup r_2 \BNFALT & \text{Regular Path} \\
             &   & r_1 \cap r_2 \BNFALT r_1 \cdot r_2 \BNFALT \NOT r \BNFALT r^* \\
  \end{array} \]%

  \end{minipage}
  \vspace{-.4em}
  \caption{\sysname core syntax.}
  \label{fig:syntax}
  \vspace{-0.8em}
\end{figure}%

@@

\begin{figure*}[t!]
  \begin{minipage}[t]{.4\linewidth}
  \hdr{Policy Automata}{}
  \vspace{1em}
    \begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=2cm,minimum size=.5cm]
      \state{0}{$0$}{              }{}
      \state{1}{$1$}{right of=0}{accepting}
      \transition{0}{0}{out}{loop above}
      \transition{0}{1}{N1}{}
      \transition{1}{1}{in}{loop above}
      \node at (1,-1){Rank 1 DFA};
    \end{tikzpicture}
    ~~~~~
    \begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=2cm,minimum size=.5cm]
      \state{0}{$0$}{              }{}
      \state{1}{$1$}{right of=0}{accepting}
      \transition{0}{0}{out}{loop above}
      \transition{0}{1}{N2}{}
      \transition{1}{1}{in}{loop above}
      \node at (1,-1){Rank 2 DFA};
    \end{tikzpicture}%
  \end{minipage}
  %
  ~~~~~~
  %
  \begin{minipage}[t]{.3\linewidth}
    \hdr{Concrete Topology}{}
    \vspace{1em}
    \includegraphics[width=.62\columnwidth]{figures/topology-con}
  \end{minipage}%
  %
  \begin{minipage}[t]{.3\linewidth}
    \hdr{Abstract Topology}{}
    \vspace{1em}
    \includegraphics[width=.505\columnwidth]{figures/topology-abs}
  \end{minipage}%

  \vspace{1em}
  \begin{minipage}[t]{.6\linewidth}
    \hdr{Concrete Product Graph}{}
    \includegraphics[width=.8\columnwidth]{figures/pg-con}
  \end{minipage}%
  \begin{minipage}[t]{.4\linewidth}
  \hdr{Abstract Product Graph}{}
    \includegraphics[width=.795\columnwidth]{figures/pg-abs}

@@

  \hrulefill
  \vspace*{-.4em}%
  \caption{Product Graph construction for policy $\True ~\Path~ \CD{\Exit(\text{N1}} ~\Prefer~ \CD{\text{N2})}$.}
  \label{fig:example-compilation}
  \vspace{-.8em}
\end{figure*}

@@

%After converting the routing policy to the regular expression based language,
We are now ready to generate the PG. Intuitively, the PG captures topology and routing constraints by ``intersecting" the finite automata associated with the policy regular expressions with the graph structure of the topology.
%The PG compactly encodes all paths that protocol messages should be passed along through the network in order to guarantee that the BGP-speaking devices find policy-compliant paths dynamically.
%
%Although \sysname policies describe the routes that traffic should take through the network, network protocols (including BGP) typically disseminate router information in the opposite direction. For example, a router Y will tell its neighbor X that it has a route to the destination meaning that X can send traffic through Y. For this reason, we start by compiling
%
Because there can be a separate routing policy for each predicate $t$ in the policy, we construct one PG for each predicate.

For each regular path constraint $r_i$ from $r_1 ~\Prefer~ ... ~\Prefer~ r_k$, we construct a DFA for the reverse of $r_i$. A DFA for $r_i$ is defined as a tuple ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma$ is the set of topology locations (\IE, routers or roles), $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
The PG for a graph $G=(V,E)$ is a tuple ($G'$, $start$, \Pref) where $G' = (V',E')$ has
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon V' \times V'$,
a unique starting vertex $start$,
and a ranking function $\Pref \colon V' \rightarrow 2^{\set{1, \dots, j}}$, mapping nodes in the PG to a set of path ranks.

The PG is constructed by adding an edge from state $m = (l_m, q_{m_1}, \dots, q_{m_k})$ to $n = (l_n, q_{n_1}, \dots, q_{n_k})$ whenever $\sigma_i(q_{m_i}, l_n) = q_{n_i}$ for each $i$ and $(l_m,l_n) \in E$.
%
We add edges from the $\mathit{start}$ node to any $m = (l, q_{m_1}, \dots, q_{m_k})$ when $\sigma_i(q_{0_i}, l) = q_{m_i}$ for each $i$.
%
The ranking function $\Pref(m)$ denotes the rank of paths through the PG ending at node $m$ and is defined as $\Pref(m) = \set{i~\vert~q_{m_i} \in F_i}$.
%
Finally, we write $\Topo{m} = l$ to extract the topology location from a PG node, when $m = (l, q_{m_1}, \dots, q_{m_k}) \in V'$. For the remainder of the paper, we use the term location to refer to both a router in the case of a concrete topology/PG, or a role in the case of an abstract topology/PG.

@@

  \hrule
  \vspace*{.4em}
  \caption{Abstract k-disjoint path analysis inference rules.}
  \label{fig:inference-rules}
  \vspace{-.8em}
\end{figure*}

@@

Figure~\ref{fig:inference-rules} displays the collection of rules used to infer facts about disjoint paths.
Each rule is read from bottom to top. The label on the bottom left is a known fact. We use $L$ to represent a rule that is parametric over the label ($S$ or $A$). Labels on other nodes correspond to facts learned after applying the rule. The box shows the conditions that must be valid, given the abstraction constraints, for the rule to apply. 
%Validity ensures that the condition holds for all concrete instantiations of the topology.

@@

@@

The abstract disjoint path analysis starts from a fixed source location $src$ and repeatedly tries to apply every inference rule from Figure~\ref{fig:inference-rules} until it reaches a fixed point. The algorithm applies an inference rule when the rule's condition is valid given the abstract topology constraints. Because the inference rules may continue to yield larger and larger symbolic expressions, we make the following observation to ensure termination: for any invariant learned of the form $L(j,k)$, it is sound to instead infer $L(j',k')$ if $j' \leq j$ and $k' \leq k$. Therefore, for each inference $L(j,k)$ we minimize the symbolic expressions for $j$ and $k$ subject to the topology constraints using the optimizing SMT solver $\nu$Z~\cite{z3opt}.

At a higher level, what is happening is that each inference rule is attempting to learn the maximum possible fault tolerance information possible as a function of the symbolic inputs, and $\nu$Z~\cite{z3opt} will minimize this maximum by accounting for all possible topologies that meet the abstraction. Facts learned with $j=0$ or $k=0$ are discarded.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figures/analysis}
  \end{center}
  \vspace{-1em}
  \caption{Abstract disjoint path analysis for global prefixes. \label{fig:compilation-times}}
  \label{fig:example-inference}
  \vspace{-.8em}
\end{figure}

@@

At a high level, the translation from the PG representation to per-device templates that run the distributed BGP protocol involves the following steps: (1) Every BGP message is tagged to record the state of the product graph as messages are passed between routers. The tags allow BGP to only search for valid paths by dropping messages that do not correspond to any edge in the PG. (2) To ensure that BGP always finds the \emph{best} paths in the network, we must rank route advertisements for each router locally such that, under this ranking function, the network as a whole satisfies the \sysname policy's end-to-end network preferences.

We introduce a simple, vendor-independent BGP configuration language called \mbgp and describe a compilation function that uses the inferred preference ordering to generate concrete configurations from a concrete topology, or templates from an abstract topology. Templates can then be soundly instantiated for a concrete network by replacing instances of abstract neighbors with the union of all concrete neighbors under the inverse homomorphism, and by replacing prefix template variables with separate entries for each concrete prefix provided by a context. We prove that concretization and compilation commute. By modifying compilation slightly we can further ensure that every configuration depends only on its immediate neighbors. This guarantees that any change made to the concrete topology will result in a minimal number of changes to the configurations.

@@

The BGP routing protocol standard allows for the use of \emph{community tags}, 32-bit integer tags that can be arbitrarily attached to, or removed from routing advertisements. 
Community tags serve as a simple form of \emph{history}, and operators routinely use such tags to implement policy. For example, operators might record where an advertisment enters the network using community tags and then conditionally block the export of the advertisment later to other peers to prevent becoming a transit point.

@@

For example, in the concrete PG in Figure~\ref{fig:example-compilation}, router S1 appears in two PG nodes: $(S1,1,0)$ and $(S1,0,1)$. These nodes have two peers that can send them messages in the PG: $(N1,0,0)$ and $(N2,0,0)$. Therefore, S1 will allow advertisements from both peer $N1$ and peer $N2$. If S1 uses the path advertised by neighbor N1, then it will add the tag $(1,0)$ to the advertisement before sending this to its neighbors A1 and A2. If S1 uses a path advertised from neighbor N2, then it will add the tag $(0,1)$ instead. Similarly, router A1 appears in two separate PG nodes. It will allow advertisements from S1 if they either have the $(1,0)$ tag or the $(0,1)$ tag attached. In either case, it will not modify the tag before exporting the route to its neighbors.

@@

  \vspace{1em}
  \hrulefill%
  \vspace{-.8em}
  \caption{mBGP syntax (left), and compilation from product graphs (right).}
  \label{fig:concretization}
  \vspace{-.8em}
\end{figure*}%

@@

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=\columnwidth]{figures/configs}
  \end{center}
  \vspace{-1em}
  \caption{Spine template and concrete configurations (left), and evolution-friendly templates (right).}
  \label{fig:bgp-configs}
  \vspace{-.6em}
\end{figure}

@@

The idea then is to search for such a device-local preference function for each router that totally orders advertisements from different neighbors (possibly with different tags). For example, to satisfy the policy that leaving through N1 is preferred to leaving through N2, S1 should prefer a message it hears from N1 over N2. Similarly, A1 should prefer routes tagged with $(1,0)$ over those with $(0,1)$. Intuitively, S1 should prefer a message from N1 because it results in an accepting state for automata 1, which indicates a better path. Further, it can result in downstream routers such as A1 using a path that is accepting for automata 1 as well.

@@

\begin{defn}
Let $m \geq_{rank} n$ be a relation over PG vertices that holds iff $\min (\Pref(m)) \geq \min (\Pref(n))$ or $\Pref(n) = \emptyset$. Intuitively $m \geq_{rank} n$ means that paths ending at node $m$ have higher automata rank and are thus no better than paths ending at $n$.
%
\end{defn}
\noindent
%

@@

Figure~\ref{fig:bgp-configs} (left) shows part of the generated \mbgp configuration for spine routers for both the concrete and abstract policies. For brevity, we using the symbol ($*$) to denote the set of all neighbors and omit tags when irrelevant.
%
For prefix $\True$, the spine routers will match advertisements from peer N1 and N2. The match for N1 is preferred since it has a higher BGP local-preference attribute (110). If an advertisement from N1 is choosen, the spine attaches the community tag $(1,0)$ before sending the route to all its peers ($*$). If an advertisement is only available from the backup N2, then it attaches the tag $(0,1)$ instead.
%
The template configuration matches any global prefix {\small \$GP} from any internal peer and re-advertises the route to all of its peers. For any local prefix it will allow an advertisement from any internal peer, and re-advertise the route to only other internal peers. The concrete configurations for S1 and S2 obtained from compilation for the concerete PG from Figure~\ref{fig:example-compilation} have a similar structure for each local and global prefix where local routes are reflected downward, while global routes are advertised to all peers.

@@

Suppose an operator wants to expand the concrete data center from Figure~\ref{fig:example} by adding an additional ToR router to the TG role. Per the network routing policy, the new router will advertise any owned prefixes provided by looking up {\small \$GP} in $\Gamma$. Because the new topology matches the abstraction, the compiled templates will remain the same. However, in the spine configurations, the match on the global prefix template variable {\small \$GP} must be expanded when concretizing the template to include the new prefixes added by the ToR. Unfortunately, this small change to the topology results in a change to every single spine configuration.

@@

\sysname is implemented in around 9500 lines of F\# code. The compiler accepts both concrete and abstract topologies and will generate evolution-friendly configurations for Cisco and Quagga~\cite{quagga} routers by default. The fault-tolerance analysis uses $\nu$Z~\cite{z3opt} to both test validity and minimize variables subject to the topology constraints. Since the analysis typically calls the SMT solver many times with relatively small optimization problems, we use a timeout of 200ms.
